\Large\bf FlexibleSUSY --- A spectrum generator generator for supersymmetric models

### 2014-06-06 doc/paper.tex
Dominik's suggestions on sections 3-4

\section{Overview of the Program}
\label{sec:Program}

@@

@@

The large variety of supersymmetric models makes it likely that the
user wants to modify the generated spectrum generator source code or
reuses components in his own program. \fs uses C++ object orientation
features to modularize the source code in order to make it easy for
the user to modify, reuse, replace or extend the individual
components.  An important application of this concept are the boundary
conditions: The boundary conditions solver provides a plugin mechanism
via a common \code{Constraint} interface, which allows a user to
exchange or add boundary conditions at any scale.  Alternatively, if
one already has an existing code for some other purpose and simply
wishes to improve upon this by adding, e.g., new RGEs or self-energies
the modular framework makes this straightforward by adding or
interfacing with the appropriate routine.  A possible application of
reusing the \fs's generated classes would be to further improve the
precision of the CE$_6$SSM spectrum generator presented in
\cite{Athron:2009bs,Athron:2012pw} by adding further one-loop
self-energies and two-loop $\beta$-functions.

@@

Furthermore, the standard algorithm which solves the user-defined
boundary value problem via a fixed-point iteration is not guaranteed
to converge in all regions of the model parameter space.  Therefore,
\fs has been intentionally designed to allow for alternative solvers
to search for solutions in such critical parameter regions.  A
subsequent release with an alternative lattice solver is already
planned.

@@

\section{Download and compilation}
\label{sec:download}

@@

\subsection{Quick Start}

@@

\subsection{Spectrum generators for alternative models}

\fs already comes with plenty of predefined models: the CMSSM (simply
called \code{MSSM}), the $Z_3$-symmetric NMSSM (called \code{NMSSM})
\cite{Ellwanger:2009dp}, $Z_3$-violating NMSSM (\code{SMSSM})
\cite{Ellwanger:2009dp}, the USSM (\code{UMSSM})
\cite{Fayet:1977yc,Cvetic:1997ky,Erler:2002pr}, the NUHM-\ESSM
(\code{E6SSM}) \cite{King:2005jy,Athron:2007en}, the right-handed
neutrino extended MSSM (\code{MSSMRHN}) and the NUHM-MSSM
(\code{NUHMSSM}).  See the content of \code{model_files/} for all
predefined model files.  A spectrum generator for the $Z_3$-symmetric
NMSSM for example can be generated like this:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=NMSSM
$ ./configure --with-models=NMSSM
$ make
\end{lstlisting}%% $
%
This NMSSM variant unifies all soft-breaking trilinear scalar
couplings at the GUT scale.  In order to relax this constraint and use
a separate value for $A_\lambda$ at the GUT scale one can edit the
model file \code{models/NMSSM/FlexibleSUSY.m} and change the lines
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

@@

\subsection{Adapting model files}

@@

\subsubsection{Integrating custom-built C++ components}

@@

>>>>

\section{Overview of the program and design goals}
\label{sec:Program}

@@

The large variety of supersymmetric models and potential
investigations makes it likely that the user wants to modify the
generated spectrum generator source code or reuse components in
further programs.  \fs offers two levels to influence the code: (i) On
the \mathematica model file level the model itself or GUT/weak scale
boundary conditions as well as input and output parameters can be
controlled (see \secref{sec:quick-start-alternative-models} and
\secref{sec:adapting-model-files} for examples).  (ii) In particular
\fs uses C++ object orientation features to modularize the source code
in order to make it easy for the user to modify, reuse, replace or
extend the individual components (see \secref{sec:adapting-cpp-code}
for examples).  An important application of this concept are the
boundary conditions, for which the C++ level offers a wider range of
possibilities.  The boundary conditions solver provides a plugin
mechanism via a common \code{Constraint} interface, which allows a
user to exchange or add boundary conditions at any scale.  An
elaborate example of a tower of effective field theories and multiple
matching scales is presented in \secref{sec:tower construction}.
Alternatively, the modular structure makes it straightforward to take
\fs generated code for e.g.\ RGEs or self-energies and reuse it in an
existing code for some other purpose.  Conversely, it is also easy to
include code from elsewhere into the spectrum generator.  For an
example see \secref{sec:integrating-custom-built}.

@@

@@

Furthermore, the standard algorithm which solves the user-defined
boundary value problem via a fixed-point iteration is not guaranteed
to converge in all regions of the model parameter space.  Therefore,
\fs has been intentionally designed to allow for alternative solvers
to search for solutions in such critical parameter regions.  A
subsequent release with an alternative solver is already planned.

@@

\section{Quick start}
\label{sec:download}

@@

\subsection{Download and generating a first spectrum generator}
\label{sec:quick-start-cmssm}

@@

\subsection{Spectrum generators for alternative models}
\label{sec:quick-start-alternative-models}

\fs already comes with plenty of predefined models: the CMSSM (simply
called \code{MSSM}), the $Z_3$-symmetric NMSSM (called \code{NMSSM})
\cite{Ellwanger:2009dp}, $Z_3$-violating NMSSM (\code{SMSSM})
\cite{Ellwanger:2009dp}, the USSM (\code{UMSSM})
\cite{Fayet:1977yc,Cvetic:1997ky,Erler:2002pr}, the NUHM-\ESSM
(\code{E6SSM}) \cite{King:2005jy,Athron:2007en}, the right-handed
neutrino extended MSSM (\code{MSSMRHN}), the NUHM-MSSM
(\code{NUHMSSM}) and the $R$-symmetric MSSM (\code{MRSSM})
\cite{Kribs:2007ac}.  See the content of \code{model_files/} for all
predefined model files.  For all these models spectrum generators can
be generated easily like for the CMSSM in
\secref{sec:quick-start-cmssm}.  The spectrum generator for the
$Z_3$-symmetric NMSSM for example can be generated like this:
%
\begin{lstlisting}[language=bash]
$ ./createmodel --name=NMSSM
$ ./configure --with-models=NMSSM
$ make
\end{lstlisting}%% $
%
One of the design goals is modularity and the possibility to easily
construct custom spectrum generators.  The details of the
customization can be found in Sections
\ref{sec:modfile}--\ref{sec:Flexible}.  As a simple example consider
the NMSSM.  The NMSSM variant above unifies all soft-breaking
trilinear scalar couplings at the GUT scale.  In order to relax this
constraint and use a separate value for $A_\lambda$ at the GUT scale
one can edit the model file \code{models/NMSSM/FlexibleSUSY.m} and
change the lines
%
\begin{lstlisting}[language=Mathematica]
EXTPAR = { {61, LambdaInput} };

@@

\subsection{Adapting model files}
\label{sec:adapting-model-files}

@@

\subsubsection{Integrating custom-built C++ components}
\label{sec:integrating-custom-built}

@@

%\cite{Kribs:2007ac}
\bibitem{Kribs:2007ac}
  G.~D.~Kribs, E.~Poppitz and N.~Weiner,
  %``Flavor in supersymmetry with an extended R-symmetry,''
  Phys.\ Rev.\ D {\bf 78} (2008) 055010
  [arXiv:0712.2039 [hep-ph]].
  %%CITATION = ARXIV:0712.2039;%%
  %130 citations counted in INSPIRE as of 06 Jun 2014
